---
title: "Project 1"
author: "Yilin Xie"
date: "June 12, 2020"
output: 
  rmarkdown::github_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data type (JSON) description

### What it is

JSON, short for JavaScript Object Notation, is a data exchange format. It is a data format popularized by Douglas Crockford in 2001 (<https://www.json.org/json-en.html>), which became the mainstream data format since 2005-2006. JSON comes in a text format that is completely independent of any programming language, making it an ideal data exchange language. JSON is the syntax for storing and exchanging textual information, similar to XML. 

### Where it gets used

There is no other place where JSON is broadly used more than the Web. Now the data interface basically returns JSON, and the detailed scenarios include:

   * Ajxa accesses data asynchronously
   * RPC remote call
   * The front and back end separate the data returned by the back end
   * Open API, such as Baidu, Aude and other open interfaces
   * Interface for cooperation between enterprises
   * Simplify object creation in JS

### Why it is a good way to store data

JSON was born because the details of XML integration into HTML vary from browser to browser. One drawback of XML formats, however, is the complexity of document construction and the need to transfer more bytes. In this case, JSON's portability gained traction, replacing XML as the dominant data transfer format. JSON is smaller, faster, and easier to parse than XML. The biggest difference between XML and JSON is that it is full of redundant information. Most of the time we don't need redundant information, but we just can't do without it when we need it. This is the biggest difference between XML and JSON. Json exists as a typical 20% feature that addresses 80% of the requirements. In other words, the fly swatter is found to be more handy where a cannon would have to be used against mosquitoes. The concise and clear hierarchy makes JSON an ideal data exchange language. Easy to read and write, and also easy to machine parsing and generation, and effectively improve the network transmission efficiency.

## Discussion of relevant R packages

Primaryly, there are three packages used for reading JSON data into R, which are `rjson`, `RJSONIO`, and `jsonlite`.

### rjson

`rjson` was first implemented for R in 2007 by Alex Couture-Beil. It allows R users to convert JSON objects into R object and vice-verse. There are three functions available under `rjson` package: *fromJSON*, *toJSON* and *newJSONParser*.

### RJSONIO

`RJSONIO` was created in 2010 by Duncan Temple Lang on GitHub. It allows the R objects are inserted into the JavaScript/ECMAScript/ActionScript code, and allow R programmers reading the JSON content and transform it into R objects.  `RJSONIO` is extensible, but it does not use the S4/S3 method. It allows others to define S4 methods for different R classes/types, as well as for callers to specify different callback handlers. Unlike the rjson package, the `RJSONIO` package USES a c++ library-libjson, instead of implementing another JSON parser, which parses faster than pure interpreted R code. There are three main functions in this package :*fromJSON*, *toJSON*, and *asJSVars*.

### jsonlite

`jsonlite` started from 2013 but has been completely rewritten in recent versions. Like `RJSONIO`, it also provides functions, such as *fromJSON()* and *toJSON() *to convert between JSON data and R objects. It could also interact with web APIs, building pipelines and streaming data between R and JSON.  

### The one I’ve chosen it
I would choose `jsonlite` in to practice for several reasons:  

  * `jsonlite` provides base64_dec and base64_enc to convert between raw vectors to text while the other two packages don’t have this function.  
  * Validating strings in JSON format is provided by `RJSONIO` (isJSONValid function) and `jsonlite` (validate) while `rjson` doesn’t have.  
  * `jsonlite` also provides the capability of re-formatting JSON file into: 1). structure with indentation added from prettify, 2). file by removing all unnecessary indentation and white spaces which is actually adopted by a lot of JavaScript libraries.  
  
  
For further information, [this page](https://rstudio-pubs-static.s3.amazonaws.com/31702_9c22e3d1a0c44968a4a1f9656f1800ab.html) gives readers a brief comparison between three packages.

## Functions to query the API

```{r message=FALSE, warning=FALSE}
## If you need, please download:
## xfun::session_info('DT')
library(DT)
library(tidyverse)
library(httr)
library(jsonlite)
```

### franchise

```{r message=FALSE}
franchise_url <- "https://records.nhl.com/site/api/franchise" 
franchise <- GET(franchise_url) %>% content("text") %>% fromJSON(flatten=TRUE)
```
```{r}
head(franchise$data, 
             caption = "Table 1.Franchise") ## assign the name of table by 'caption = ""'
```


### franchise team

```{r message=FALSE}
team_url <- "https://records.nhl.com/site/api/franchise-team-totals"
team <- GET(team_url) %>% content("text") %>% fromJSON(flatten=TRUE)
```
```{r}
head(team$data, 
             caption = "Table 2.Total stats for every franchise")
```

### franchise season records

You can assign a certain value to specify the franchise in column **franchiseId** and **franchiseName**. This is applicable to the following three tables.
```{r message=FALSE}
## combine all JSON files so that make it possible to apply interactive filtering
season<-data.frame()
for (i in 1:40){
  season_url <- paste("https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId",i,collapse="",sep = "=")
  season0 <- GET(season_url) %>% content("text") %>% fromJSON(flatten=TRUE)
  season <- rbind(season, season0$data)
}
## put the two columns first for the user to find easily
season1 <- subset(season, select=-c(franchiseId, franchiseName))
col <- colnames(season1)
new_col <- c(col[1], "franchiseId","franchiseName", col[2:length(col)]) ## specify the order of columns
season2 <- season[new_col] ## arrange the columns in the specified order
```
```{r}
## interactive table for users to specify the franchise
head(season2, 
          filter = 'top', 
          rownames = FALSE,
          caption = 'Table 3.Season records for a specific franchise')
```

### franchise goalie records

```{r message=FALSE}
goalie<-data.frame()
for (i in 1:40){
  goalie_url <- paste("https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId",i,collapse="",sep = "=")
  goalie0 <- GET(goalie_url) %>% content("text") %>% fromJSON(flatten=TRUE)
  goalie <- rbind(goalie, goalie0$data)
}
goalie1 <- subset(goalie, select=-c(franchiseId, franchiseName))
col <- colnames(goalie1)
new_col <- c(col[1], "franchiseId", "franchiseName", col[2:length(col)])
goalie2 <- goalie[new_col]
```
```{r}
head(goalie2, 
          filter = 'top', 
          rownames = FALSE,
          caption = 'Table 4.Goalie records for a specified franchise')
```


### franchise skater records
```{r message=FALSE}
skater<-data.frame()
for (i in 1:40){
  skater_url <- paste("https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId",i,collapse="",sep = "=")
  skater0 <- GET(skater_url) %>% content("text") %>% fromJSON(flatten=TRUE)
  skater <- rbind(skater, skater0$data)
}
skater1 <- subset(skater, select=-c(franchiseId, franchiseName))
col <- colnames(skater1)
new_col <- c(col[1], "franchiseId", "franchiseName", col[2:length(col)])
skater2 <- skater[new_col]
```
```{r warning=FALSE}
head(skater2, 
          filter = 'top',
          rownames = FALSE,
          caption = 'Table 5.Skater records for the specified franchise')
```

## Data analysis
### Creation of relevant new variable
This part is done with dataset **franchise team** which contains total stats for every franchise.
```{r message=FALSE, warning=FALSE}
## ratio and difference of goals for and goals against
team <- mutate(team$data, ratio=goalsFor/goalsAgainst)
team <- mutate(team, diff=goalsFor-goalsAgainst)
## home ice advantage
team <- mutate(team, homeAd=homeWins/wins)
## home ice disadvantage
team <- mutate(team, homeDis=homeLosses/losses)
## win rate
team <- mutate(team, winRate=wins/gamesPlayed)
```

### Calculation of relevant numeric & graphical summaries
```{r message=FALSE, warning=FALSE}
library(ggplot2)
```

#### Some contingency tables and numeric summaries
```{r}
## a contingency table of active goalies of five franchises
goalie2 %>% filter(franchiseId<=5) %>% with(table(activePlayer,franchiseName)) ## use 'table()' to create a contingency table
```

The contingency table includes the teams with ID 1 to 5, Montréal Canadiens, Montreal Wanderers, St. Louis Eagles, Hamilton Tigers, Toronto Maple Leafs, respectively. This table counts active goalie players of each franchise. Three of franchises have no active players, and the total munber of goalie players is small. Toronto Maple Leafs, besides the total number of players is higher than that of Montreal Canadiens, the proportion of active players is higher.


```{r}
## comparison of goalies of Montréal Canadiens and Los Angeles Kings by numeric summaries of two franchises
MC <- goalie2 %>% filter(franchiseId == "1") %>% select(activePlayer, gamesPlayed, losses, seasons, wins, mostGoalsAgainstOneGame, mostSavesOneGame, mostShotsAgainstOneGame, mostWinsOneSeason) 
apply(as.matrix(MC), 2,
      function(x)
        c(n = sum(!is.na(x)),
        mean = mean(x, na.rm=TRUE),
        sd = sd(x, na.rm=TRUE),
        min = min(x, na.rm=TRUE),
        max = max(x, na.rm=TRUE),
        range = max(x, na.rm=TRUE)-min(x, na.rm=TRUE))) %>% knitr::kable(caption = "Table 7a. Summary of goalies of Montréal Canadiens")
LAK <- goalie2 %>% filter(franchiseId == "14") %>% select(activePlayer, gamesPlayed, losses, seasons, wins, mostGoalsAgainstOneGame, mostSavesOneGame, mostShotsAgainstOneGame, mostWinsOneSeason) 
apply(as.matrix(LAK), 2,
      function(x)
        c(n = sum(!is.na(x)),
        mean = mean(x, na.rm=TRUE),
        sd = sd(x, na.rm=TRUE),
        min = min(x, na.rm=TRUE),
        max = max(x, na.rm=TRUE),
        range = max(x, na.rm=TRUE)-min(x, na.rm=TRUE))) %>% knitr::kable(caption = "Table 7b. Summary of goalies of Los Angeles Kings")
```

Table 7 shows the goalie performance of the Montreal Canadiens (MC) and the Los Angeles Kings (LAK). MC has 37 goalies and LAK has 43 goalies. MC goalies have more experience, averagely participateed in 123 games, about 1.5 times as many as LAK did. The average number of wins for MC goalies is twice that of LAK goalies, but this does not mean that MC goalies perform better. It can be inferred from the *mostSavesOneGame* that MC goalies' winning rates may be due to strong teammates and poor performance compared to LAK goalies. From the perspective of SD and Range, MC goalkeeper's strength is more uneven, while LAK goalkeeper's strength is more balanced.

#### A side-by-side bar plot with coloring

```{r}
## a bar plot of skaters of each franchise
p <- ggplot(data = skater2,
            mapping = aes(x = franchiseName, fill = franchiseName )) + geom_bar() + coord_flip() + labs(x = "franchise", y = "total number of skaters", title = "Figure 1. The total number of skaters of each franchise") + theme(legend.position = "none")
p
```

Figure 1 shows the total number of skaters each franchise has. As can be seen from the figure, the size of the teams varies greatly. About half of all franchise have more than 500 skaters, with the New York Rangers leading the way with nearly 1,000. Some, like the Montreal Wanderers, Vegas Golden Knights, Phiadelphia Flyers and Hamilton Tigers, are small and pitiable. If you want to know the exact number of players for these teams, you can filter them using interaction in Table 5.

#### A side-by-side box plots with coloring

```{r warning=FALSE}
## a box plot of skater's goals of each franchise
bp <- ggplot(data = skater2,
             mapping = aes(x = franchiseName, y = log(goals), fill = franchiseName)) + geom_boxplot() + coord_flip() + labs(x = "franchise", title = "Figure 2. The total goals scored by skaters of each franchise") + theme(legend.position = "none")
bp
```

Figure 2 shows the distribution of goals scored by players on each franchise. Because the high numerical value is too discrete, we take the logarithm of goal. From the data distribution, we can find that the score dispersion degree of most franchises is similar, although there is a certain difference in scoring ability. The numbers of skaters of Montreal Wanderers and Vegas Golden Knights teams are relatively small, but the goals of the Montreal Wanderers are ridiculously low, while the goals of the Vegas Golden Knights team are at the top of all the teams, so it can be said that they are small but good.

#### A scatter plots with coloring

```{r, fig.height=7, fig.width=10} 
## a scatter plot of skater's points of each franchise
## to better present the figure, you can assign its size by ```{r, fig.height=7, fig.width=10} 
sp <- ggplot(data = skater2, 
             mapping = aes(
               x = assists,
               y = goals,
               color = franchiseName
             )) + geom_point(alpha = 0.4) + labs(title = "Figure 3.The relationship between goals and assists scored by skaters of each franchise")+ theme(legend.position = "bottom")
sp

```

Figure 3 shows the relationship between goals and assists of each skater. On the whole, the number of goals has a positive correlation with the number of assists gained, about assists =1.5*goals. But for each player's specific performance, the total number of assists and points the total difference is still relatively large. Most players have less than 300 assists and score less than 200 points.  


