---
title: "Project 1"
author: "Yilin Xie"
date: "June 12, 2020"
output: 
  rmarkdown::github_document:
    toc: yes
---

## Data type (JSON) description

### What it is

JSON, short for JavaScript Object Notation, is a data exchange format. It is a data format popularized by Douglas Crockford in 2001 (<https://www.json.org/json-en.html>), became the mainstream data format since 2005-2006. JSON comes in a text format that is completely independent of any programming language, making it an ideal data exchange language. JSON is the syntax for storing and exchanging textual information, similar to XML. 

### Where it gets used

There is no other place where JSON is broadly used more than the Web. Now the data interface basically returns JSON, and the detailed scenarios include:

   * Ajxa accesses data asynchronously
   * RPC remote call
   * The front and back end separate the data returned by the back end
   * Open API, such as Baidu, Aude and other open interfaces
   * Interface for cooperation between enterprises
   * Simplify object creation in JS

### Why it is a good way to store data

JSON was born because the details of XML integration into HTML vary from browser to browser. One drawback of XML formats, however, is the complexity of document construction and the need to transfer more bytes. In this case, JSON's portability gained traction, replacing XML as the dominant data transfer format. JSON is smaller, faster, and easier to parse than XML. The biggest difference between XML and JSON is that it is full of redundant information. Most of the time we don't need redundant information, but we just can't do without it when we need it. This is the biggest difference between XML and JSON. Json exists as a typical 20% feature that addresses 80% of the requirements. In other words, the fly swatter is found to be more handy where a cannon would have to be used against mosquitoes. The concise and clear hierarchy makes JSON an ideal data exchange language. Easy to read and write, and also easy to machine parsing and generation, and effectively improve the network transmission efficiency.

## Discussion of relevant R packages

Primaryly, there are three packages used for reading JSON data into R, which are `rjson`, `RJSONIO`, and `jsonlite`.

### rjson

`rjson` was first implemented for R in 2007 by Alex Couture-Beil. It allows R users to convert JSON objects into R object and vice-verse. There are three functions available under `rjson` package: *fromJSON*, *toJSON* and *newJSONParser*.

### RJSONIO

`RJSONIO` was created in 2010 by Duncan Temple Lang on GitHub. It allows the R objects are inserted into the JavaScript/ECMAScript/ActionScript code, and allow R programmers reading the JSON content and transform it into R objects.  `RJSONIO` is extensible, but it does not use the S4/S3 method. It allows others to define S4 methods for different R classes/types, as well as for callers to specify different callback handlers. Unlike the rjson package, the `RJSONIO` package USES a c++ library-libjson, instead of implementing another JSON parser, which parses faster than pure interpreted R code. There are three main functions in this package :*fromJSON*, *toJSON*, and *asJSVars*.

### jsonlite

`jsonlite` started from 2013 but has been completely rewritten in recent versions. Like `RJSONIO`, it also provides functions, such as *fromJSON()* and *toJSON() *to convert between JSON data and R objects. It could also interact with web APIs, building pipelines and streaming data between R and JSON.

I would choose `jsonlite` in to practice for several reasons:

  * `jsonlite` provides base64_dec and base64_enc to convert between raw vectors to text while the other two packages don’t have this function. 
  * Validating strings in JSON format is provided by `RJSONIO` (isJSONValid function) and `jsonlite` (validate) while `rjson` doesn’t have. 
  * `jsonlite` also provides the capability of re-formatting JSON file into: 1). structure with indentation added from prettify, 2). file by removing all unnecessary indentation and white spaces which is actually adopted by a lot of JavaScript libraries. 
  
  
For further information, [this page](https://rstudio-pubs-static.s3.amazonaws.com/31702_9c22e3d1a0c44968a4a1f9656f1800ab.html) gives readers a brief comparison between three packages.

## Functions to query the API

```{r message=FALSE, warning=FALSE}
## If you need, please download:
## xfun::session_info('DT')
library(DT)
library(tidyverse)
library(httr)
library(jsonlite)
```



```{r message=FALSE}
#write a function
myfunc <- function(x){
  full_url <- paste0("https://records.nhl.com/site/api", "/", x)
  #Returns id, firstSeasonId and lastSeasonId and name of every team in the history of the NHL
  if(x == "franchise"){
    franchise <- GET(full_url) %>% content("text") %>% fromJSON(flatten=TRUE)
    data <- franchise$data %>% select(id,firstSeasonId,lastSeasonId,teamCommonName)
    tbl_df(data)
  }
  #Returns Total stats for every franchise
  else if(x == "franchise-team-totals"){
    franchise <- GET(full_url) %>% content("text") %>% fromJSON(flatten=TRUE)
    data <- franchise$data %>% select(id,roadTies,roadWins)
    tbl_df(data)
  }
  #Drill-down into season records for a specific franchise
  else if (x %in% "franchise-season-records?cayenneExp=franchiseId="){
    franchiseId <- substr(x, 49, 50)
    #convert a character vector to a numeric vector
    franchiseId <- as.numeric(franchiseId)
    franchise <- GET(full_url) %>% content("text") %>% fromJSON(flatten=TRUE)
    franchise$data %>% filter(id == franchiseId)
  }
  #Goalie records for the specified franchise
  else if (x %in% "franchise-goalie-records?cayenneExp=franchiseId="){
    franchiseId <- substr(x, 49, 52)
    #convert a character vector to a numeric vector
    franchiseId <- as.numeric(franchiseId)
    franchise <- GET(full_url) %>% content("text") %>% fromJSON(flatten=TRUE)
    franchise$data %>% filter(id == franchiseId)
  }
  #Skater records, same interaction as goalie endpoint
  else if (x %in% "franchise-skater-records?cayenneExp=franchiseId="){
    franchiseId <- substr(x, 49, 53)
    #convert a character vector to a numeric vector
    franchiseId <- as.numeric(franchiseId)
    franchise <- GET(full_url) %>% content("text") %>% fromJSON(flatten=TRUE)
    franchise$data %>% filter(id == franchiseId)
  }
}
myfunc("franchise")
myfunc("franchise-team-totals")
myfunc("franchise-season-records?cayenneExp=franchiseId=10")
myfunc("franchise-goalie-records?cayenneExp=franchiseId=235")
myfunc("franchise-skater-records?cayenneExp=franchiseId=16891")
#attributes(franchise)
#franchise$data
```


